Simulation and visualization framework for decision making of climate mitigation \vspace{4pt} and adaptation strategies in new and existing urban development

### 2021-01-31 manuscript_wesbrook-bigmacc_BS2021.tex

The resulting simulation database can be used to assess decarbonization pathways that align with development trajectories and/or different degrees of global warming. The difficulty though is in interpreting a large quantity of simulation data for its salience in the decarbonization and adaptation space. Several (\citep{}) have proposed marginal abatement cost curves (MACCs) as a useful metric to compare strategies, while \citet{kesicki_marginal_nodate} points out that their use should be tempered with caution related to the lack of interaction of strategies in typical MACCs, a point taken up in \citet{rysanek MACC} through an underlying sequential and exhaustive simulation framework. Applying this same framework to the task of determining optimal decarbonization pathways through mitigation strategies can result in a viable set of MACCs for planing and policy support.

@@

Wesbrook Village and the neighbouring Stadium Road are new communities that expected to reach 715,925 sqm. of residential floor space and 10,000 sqm. of non-residential floor space by 2040. It is the part of a larger community plan led by the development arm of The University of British Columbia (UBC) \citep{wesbrook plan; stadium plan}. The baseline for this study is taken at the end of 2020, where the development has completed its fifteenth year. The development, referred to throughout the study as \textit{Wesbrook}, is a compelling case study as it can be considered a model of medium-density new development for BC and North America at large. Building code for the community is modeled after the aspirational provincial code made more strict through a continuous review with the most recent version taking effect in January 2021 and targeting net-zero energy ready (NZER) by 2032; NZER buildings are built such that with the addition of solar panels or renewable technologies they would consume only as much energy as they produce. The energy system is a mixture of grid-connected systems, in building natural gas boilers, and a district heating system that supplies hydronic baseboard and floor heating systems as well as domestic hot water. This district heating system, described in \citet{McCarty 2021}, is designed to migrate from natural gas boilers to heat pumps fed by waste heat from a nearby particle accelerator facility that is trimmed up by peaking natural gas boilers. Cooling services have yet to considered in spec-residential development in BC as the climate has historically allowed comfort to be achieved in the peak of summer through natural ventilation; research however indicates that this is quickly becoming an obsolete design standard due to climate change \citep{ML cooling paper}. The intention with the code development for this community will however require cooling in buildings soon.

\subsubsection*{Retrofit and New Development Trajectory}
The majority of Wesbrook has been completed as of 2020, with 50\% of the total expected build-out of 725,925 sqm. remaining between now and the mid-21st century. As indicated in Figure \ref{fig:wesbrook_base} the yet-to-be-completed development will be assessed in the model through the future periods. While in less centrally planned and developed communities tackling development uncertainty would require a more thorough economic model, the Wesbrook planning and development scheme is well laid out in master-planning documents \citep{Wesbrook masterplan}. The simulation framework is built from a baseline in 2020 of the completed development up to 2020. Two more time periods were simulated, the first in 2050 and then in 2080. The base-case for each time period assumes that the stock of buildings that were simulated in the previous time-step (existing) were retrofitted with similarly performing technologies and components. New buildings are however built to a performance profile modeled after the changes in the municipalities building performance code (\citep{REAP}). While it is expected that the future will hold more advanced technologies, code development is capped at the introduction of the NZER requirements.

@@

\subsection*{Mitigation and adaptation decision tools}
\citet{Rysanek MACC} expanded on the use of marginal abatement cost curves as a tool for planning retrofit pathways in single building optimization. The authors commentary was that by failing to consider technical and economic interactions MACCs such as those proposed by \citet{Mckinnsey} were more \textit{illustrative} than planning tools. Without accounting for the additionality of measures many MACCs fall into this grouping and limit their overall effectiveness. \citet{kesicki} acknowledges temporal interactions as an additional dimension that need to be considered when constructing MACCs. Traditionally MACCs were developed for annual-planning and more recently, as ambition has grown, so too have their temporal scales. These interactions can be summarized as understanding how investments made in the past may affect the outcome of the MACC, similar to the issue of technical and economic interactions. In planning buildings or infrastructure with MACCs this becomes a significant feature to consider as selecting investments to implement, one typically does so along an investment pathway.

It is also of particular importance when developing infrastructure or buildings to consider the impact of a changing climate. Adaptation is prevalent in designing buildings in proximity to rising seas, adaptation ranges from hard to soft measures such as levies and dams to wetland restoration \citep{IPCC WGIIAR5 CH 15}. Evidence is pointing to the need to consider other impacts, such as overheating risk due to generally warmer days, nights with less cooling capacity, and more intense heat waves \citep{Lomas and Porrit; Rysanek ML Paper}. Measuring adaptation is a contentious field, as what defines successful adaptation is not as simple to calculate as net carbon emissions \citep{IPCC WGIIAR5 CH 14}. \citet{Brooks et al.} suggests several general criteria by which successful adaptation can be assessed: feasibility, efficacy/effectiveness, efficiency, acceptability/legitimacy, equity, and sustainability. The same authors also discuss the concept of Maladaptation, or the inadvertent increase in vulnerability to climate change as a result of development, which has equal importance to adaptation planning of buildings. For all criteria though, measures begin with assessing vulnerability of a subject to a stressor or collection of stressors. In the case of coastal development, a building's adaptive capacity and success can be measured by its vulnerability to increasingly intense sea level rise and storm surge. In the case of this study the stressor is heat gain and the vulnerability of a building to overheating.

>>>>

The resulting simulation database can be used to assess decarbonization pathways that align with development trajectories and/or different degrees of global warming. The difficulty though is in interpreting a large quantity of simulation data for its salience in the decarbonization and adaptation space. Several (\citep{afshari_life-cycle_2014,ibrahim_methodology_2016}) have proposed marginal abatement cost curves (MACCs) as a useful metric to compare strategies, while \citet{kesicki_marginal_nodate} points out that their use should be tempered with caution related to the lack of interaction of strategies in typical MACCs, a point taken up in \citet{rysanek_using_2013} through an underlying sequential and exhaustive simulation framework. Applying this same framework to the task of determining optimal decarbonization pathways through mitigation strategies can result in a viable set of MACCs for planing and policy support.

@@

Wesbrook Village and the neighbouring Stadium Road are new communities that expected to reach 715,925 sqm. of residential floor space and 10,000 sqm. of non-residential floor space by 2040. It is the part of a larger community plan led by the development arm of The University of British Columbia (UBC) \citep{ubc_planning_ubc_2020,ubc_planning_ubc_2020_SR}. The baseline for this study is taken at the end of 2020, where the development has completed its fifteenth year. The development, referred to throughout the study as \textit{Wesbrook}, is a compelling case study as it can be considered a model of medium-density new development for BC and North America at large. Building code for the community is modeled after the aspirational provincial code made more strict through a continuous review with the most recent version taking effect in January 2021 and targeting net-zero energy ready (NZER) by 2032; NZER buildings are built such that with the addition of solar panels or renewable technologies they would consume only as much energy as they produce. The energy system is a mixture of grid-connected systems, in building natural gas boilers, and a district heating system that supplies hydronic baseboard and floor heating systems as well as domestic hot water. This district heating system, described in \citet{mccarty_accepted_2020}, is designed to migrate from natural gas boilers to heat pumps fed by waste heat from a nearby particle accelerator facility that is trimmed up by peaking natural gas boilers. Cooling services have yet to considered in spec-residential development in BC as the climate has historically allowed comfort to be achieved in the peak of summer through natural ventilation; research however indicates that this is quickly becoming an obsolete design standard due to climate change \citep{rysanek_forecasting_2021}. The intention with the code development for this community will however require cooling in buildings soon.

\subsubsection*{Retrofit and New Development Trajectory}
The majority of Wesbrook has been completed as of 2020, with 50\% of the total expected build-out of 725,925 sqm. remaining between now and the mid-21st century. As indicated in Figure \ref{fig:wesbrook_base} the yet-to-be-completed development will be assessed in the model through the future periods. While in less centrally planned and developed communities tackling development uncertainty would require a more thorough economic model, the Wesbrook planning and development scheme is well laid out in master-planning documents \citep{noauthor_ubc_2020_wb}. The simulation framework is built from a baseline in 2020 of the completed development up to 2020. Two more time periods were simulated, the first in 2050 and then in 2080. The base-case for each time period assumes that the stock of buildings that were simulated in the previous time-step (existing) were retrofitted with similarly performing technologies and components. New buildings are however built to a performance profile modeled after the changes in the municipalities building performance code (\citep{noauthor_residential_2021}). While it is expected that the future will hold more advanced technologies, code development is capped at the introduction of the NZER requirements.

@@

\subsection*{Mitigation and adaptation decision tools}
\citet{rysanek_using_2013} expanded on the use of marginal abatement cost curves as a tool for planning retrofit pathways in single building optimization. The authors commentary was that by failing to consider technical and economic interactions MACCs such as those proposed by \citet{mckinsey__co_climate_2010} were more \textit{illustrative} than planning tools. Without accounting for the additionality of measures many MACCs fall into this grouping and limit their overall effectiveness. \citet{kesicki_marginal_nodate} acknowledges temporal interactions as an additional dimension that need to be considered when constructing MACCs. Traditionally MACCs were developed for annual-planning and more recently, as ambition has grown, so too have their temporal scales. These interactions can be summarized as understanding how investments made in the past may affect the outcome of the MACC, similar to the issue of technical and economic interactions. In planning buildings or infrastructure with MACCs this becomes a significant feature to consider as selecting investments to implement, one typically does so along an investment pathway.

It is also of particular importance when developing infrastructure or buildings to consider the impact of a changing climate. Adaptation is prevalent in designing buildings in proximity to rising seas, adaptation ranges from hard to soft measures such as levies and dams to wetland restoration \citep{}. Evidence is pointing to the need to consider other impacts, such as overheating risk due to generally warmer days, nights with less cooling capacity, and more intense heat waves \citep{Lomas and Porrit; Rysanek ML Paper}. Measuring adaptation is a contentious field, as what defines successful adaptation is not as simple to calculate as net carbon emissions \citep{IPCC WGIIAR5 CH 14}. \citet{Brooks et al.} suggests several general criteria by which successful adaptation can be assessed: feasibility, efficacy/effectiveness, efficiency, acceptability/legitimacy, equity, and sustainability. The same authors also discuss the concept of Maladaptation, or the inadvertent increase in vulnerability to climate change as a result of development, which has equal importance to adaptation planning of buildings. For all criteria though, measures begin with assessing vulnerability of a subject to a stressor or collection of stressors. In the case of coastal development, a building's adaptive capacity and success can be measured by its vulnerability to increasingly intense sea level rise and storm surge. In the case of this study the stressor is heat gain and the vulnerability of a building to overheating.
